---
title: "hcol_courses"
author: "JBR"
date: "4/20/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)


library(broom)
library(dplyr)
library(tidyverse)
library(tidyr)
library(gt)
library(ggplot2)
library(infer)
library(forcats)
library(purrr)
library(skimr)
library(readr)
library(readxl)
library(janitor)
library(ggthemes)


spring_2016 <- read_xlsx("ENROLLMENT/spring_2016.xlsx")%>%
  clean_names()

spring_2017 <- read_xlsx("ENROLLMENT/spring_2017.xlsx")%>%
  clean_names()

spring_2018 <- read_xlsx("ENROLLMENT/spring_2018.xlsx")%>%
  clean_names()

spring_2019<- read_xlsx("ENROLLMENT/spring_2019.xlsx", skip = 3)%>%
  clean_names()

spring_2020 <- read_xlsx("ENROLLMENT/spring_2020.xlsx")%>%
  clean_names()

fall_2016 <- read_xlsx("ENROLLMENT/fall_2016.xlsx")%>%
  clean_names()

fall_2017 <- read_xlsx("ENROLLMENT/fall_2017.xlsx", skip = 3)%>%
  clean_names()

fall_2018 <- read_xlsx("ENROLLMENT/fall_2018.xlsx", skip = 2)%>%
  clean_names()

fall_2019 <- read_xlsx("ENROLLMENT/fall_2019.xlsx", skip = 3)%>%
  clean_names()


```

Basic summary analysis: 
- What are the biggest courses? 
    - Enrollment?
- What are the biggest departments?
    - Enrollment?
    - Course Count? 


```{r}

# This chunk seeks to answer: What were the biggest courses for the 2019-2020 Academic Year?

ay_1920 <- spring_2020%>%
  full_join(fall_2019)%>%
  slice(1:5244)%>%
  filter(course_id != "Grand Total")

top_10_ay_1920 <- ay_1920 %>%
  arrange(desc(u_grad))%>%
  slice(1:10)

top_10_ay_1920 %>%
  select(course_id, course_title, course_name, course_department, instructor_full_name, u_grad)%>%
  gt()%>%
  tab_header(
    title= "10 Biggest Courses During AY 19-20"
  )%>%
  cols_label(course_id = "Course ID",
             course_title = "Course Title",
             course_name = "Name",
             course_department = "Department",
             instructor_full_name = "Instructor",
             u_grad = "Undergrads")


```

ADDITIONAL THINGS IT WOULD BE INTERESTING TO DO: FIND AVERAGE COURSE SIZE

```{r AVERAGE COURSE SIZE}
ay_1920 %>%
  summarize(mean_class_size = mean(u_grad))

```

OTHER THINGS IT MIGHT BE INTERESTING TO DO: CREATE A FORMULA FOR COURSE SIZE AND GENERAL EXPERIENCE, LIKE TF'S ETC



```{r VISUALIZATION OF BIG 10 ENROLLMENTS}
# This chunk seeks to visualize the % of undergrad 
# students enrolled in the "big 10" at any time. 

# summarizing for big 10 enrollment

top_10_ay_1920 %>%
  summarize(ug_big10 = sum(u_grad))

# 4689 in big 10 for ay1920

# summarizing for total enrollment 
# (not STUDENTS but STUDENTS in CLASSES 
# so will by much larger than amount of students)

ay_1920 %>%
  summarize(ay_stu = sum(u_grad))%>%
  view()

# 52389

# 4689/52389 yields 0.08950,
# so it looks like 9% 
# of the student body
# is taking or has taken  one of the 
# top 10 enrolled classes for AY19-20.

# now i'm going to create a two-level factor
# for big 10 or non big 10, so i can stack
# the bar_chart and fill by that factor

big_10_plot <- ay_1920 %>%
  arrange(desc(u_grad))%>%
  filter(course_id != "Grand Total")%>%
  mutate(big_10 = ifelse(u_grad > 266, TRUE, FALSE))%>%
  mutate(big_10 = as.factor(big_10)) 
  
# SIMPLE IS BEST
# PERCENT OF STUDENT IN BIG 10 



big_10_plot_sa <- big_10_plot %>%
  ggplot(aes(x = 1, y = u_grad, fill = big_10))+
  geom_bar(position = "stack", stat = "identity")+
  scale_x_discrete()+
  annotate("text", x = 1, y = 2400, label = "9%")+
annotate("text", x = 1, y = 27000, label = "91%")+
  xlab("")+
  ylab("Total Students")+
  labs(title = "% of Course Enrollments during AY 19-20", 
       fill = "Big 10 Class")+
  theme_fivethirtyeight()
  
# saving image to print 

ggsave("big_10_plot_sa.png",
      plot = last_plot(), path = "shiny/")
  
# saving the rds to print 
saveRDS(big_10_plot, file="big_10_plot.Rds")
  
```

```{r BIGGEST DEPARTMENTS BY ENROLLMENT GT}
# This seeks to find the biggest departments in AY19-20

# by total enrollment in classes

departments_ay1920 <- ay_1920 %>%
  group_by(course_department)%>%
  summarize(dept_enrollments = sum(u_grad))%>%
  filter(dept_enrollments != 0 & course_department != "General Education")

# rank depts by enrollment

depts_ranked_ay1920 <- departments_ay1920 %>%
  arrange(desc(dept_enrollments)) 

# create gtable
depts_ranked_ay1920%>%
  slice(1:10)%>%
  gt()%>%
    tab_header(title = "Total Amount of Enrollments in Courses by Department",
               subtitle = "AY 19-20")%>%
    cols_label(course_department = "Department",
         dept_enrollments = "Enrollments")

# SHINY
dray1920_sa <- depts_ranked_ay1920%>%
  slice(1:10)%>%
  gt()%>%
    tab_header(title = "Total Amount of Enrollments in Courses by Department",
               subtitle = "AY 19-20")%>%
    cols_label(course_department = "Department",
         dept_enrollments = "Enrollments")
ggsave("dray1920_sa.png",
      plot = last_plot(), path = "shiny/")


```


```{r BIGGEST DEPTS BY ENROLLMENT PLOT}
# graph 
depts_ranked_ay1920 %>%
  arrange(desc(dept_enrollments))%>%
  slice(1:10)%>%
 ggplot(aes(x = course_department, y = dept_enrollments, fill = course_department))+
  geom_bar(stat = "identity")+
  scale_x_discrete()+
  theme(axis.text.x = element_text(angle = 70, hjust = 1))+
  labs(title = "Total Enrollments in Top 10 Departments", fill = "Department")+
  xlab("")+
  ylab("Total Enrollments")



```

```{r SHINY APP ENROLLMENT 1920 PLOT}


dray1920_plot_enrollment_sa <- depts_ranked_ay1920 %>%
  arrange(desc(dept_enrollments))%>%
  slice(1:10)%>%
 ggplot(aes(x = course_department, y = dept_enrollments, fill = course_department))+
  geom_bar(stat = "identity")+
  scale_x_discrete()+
  theme(axis.text.x = element_text(angle = 70, hjust = 1))+
  labs(title = "Total Enrollments in Top 10 Departments",
       subtitle = "AY 2019-2020",
       fill = "Department")+
  xlab("")+
  ylab("Total Enrollments")

ggsave("dray1920_plot_enrollment_sa.png",
      plot = last_plot(), path = "shiny/")
```

**CONSIDER GGANIMATE**

```{r BIGGEST DEPTS BY COURSE COUNT GT}

dept_cc1920 <- ay_1920 %>%
  group_by(course_department)%>%
  summarize(course_count = n())%>%
  filter(course_department != "Medical Sciences")

dept_cc1920 %>%
  arrange(desc(course_count))%>%
  slice(1:15) %>%
  gt()%>%
  tab_header(
    title= "15 Biggest Departments by Course Count",
    subtitle = "AY 19-20"
  )%>%
  cols_label(
    course_department = "Department",
    course_count = "Number of Courses"
  )
```

```{r GGPLOT}

dept_cc1920 %>%
  arrange(desc(course_count))%>%
  slice(1:15)%>%
    ggplot(aes(course_department, course_count, fill = course_department))+
  geom_bar(stat = "identity")+
  geom_text(aes(label=course_count), position=position_dodge(width=0.9), vjust=-0.25)+
  scale_x_discrete()+
  theme(axis.text.x = element_text(angle = 70, hjust = 1))+
  labs(title = "Academic Departments with Most Courses", fill = "Department")+
  xlab("")+
  ylab("Total Courses")

ggsave("dept_cc1920.png",
      plot = last_plot(), path = "shiny/")

saveRDS(dept_cc1920, file="dept_cc1920.Rds")

```

**CONSIDER A GGANIMATE PLOT CORRESPONDING TO EACH YEAR**

Temporal Summary Analysis: 
- What are the fastest growing departments from the last 4 years?

```{r}
# in order to pull this off, I need to create AY 
# for the last 3 years. 
# luckily, I already have one of them (AY1920)

# creating AY_1819

ay_1819 <- spring_2019%>%
  full_join(fall_2018)%>%
  slice(1:5182)%>%
  filter(course_id != "Grand Total")

# refining departments 
departments_ay1819 <- ay_1819 %>%
  group_by(course_department)%>%
  summarize(dept_enrollments = sum(u_grad))%>%
  filter(dept_enrollments != 0 & course_department != "General Education")

# creating growth 

depts_18_to20 <- departments_ay1819 %>%
  full_join(departments_ay1920, by = "course_department", suffix = c("18", "19"))

# growth in integer and percent 

growth_1819 <- depts_18_to20 %>%
  mutate(growth_1819_pct = ((dept_enrollments19 - dept_enrollments18)/dept_enrollments18),
         growth_1819 = (dept_enrollments19 - dept_enrollments18))

# PROCEDURAL THINGS 
# 1. CREATE AY 2017-2018 DEPTS AND JOIN 
# MEAN ANNUAL CHANGE IN ENROLLMENT GRAPH
# 2. CREATE A NEW GROUPING AND MAKE 
# IT FOR COURSE COUNT

ay_1819 <- spring_2019%>%
  full_join(fall_2018)%>%
  slice(1:5182)%>%
  filter(course_id != "Grand Total")

ay_1718 <- spring_2018 %>%
  full_join(fall_2017)%>%
  filter(course_id != "Grand Total")%>%
  slice(1:5040)

# refining departments 
departments_ay1718 <- ay_1718 %>%
  group_by(course_department)%>%
  summarize(dept_enrollments = sum(u_grad))%>%
  filter(dept_enrollments != 0 & course_department != "General Education")

# creating growth 

depts_17_to20 <- departments_ay1718 %>%
  full_join(depts_18_to20, by = "course_department")%>%
  rename(dept_enrollments17 = dept_enrollments)

# growth in integer and percent 

growth_1718 <- depts_17_to20 %>%
  mutate(growth_1718_pct = ((dept_enrollments18 - dept_enrollments17)/dept_enrollments17),
         growth_1718 = (dept_enrollments18 - dept_enrollments17))

growth_17_to_19 <- growth_1718 %>%
  right_join(growth_1819)

# creating mean growth over three years

mean_growth_3y <- growth_17_to_19%>%
  filter(course_department != "Medical Sciences")%>%
  na.omit()%>%
  mutate(mean_growth_enrollment = (growth_1718 + growth_1819)/2,
            mean_growth_pct = (((growth_1819_pct + growth_1819_pct)/2)*(100)))%>%
  mutate_if(is.numeric, ~round(., 2))




```

```{r Mean Growth Percentage PLOT}

mean_growth_3y %>%
  # exclude med sciences
  arrange(desc(mean_growth_pct))%>%
  slice(1:10)%>%
  ggplot(aes(course_department, mean_growth_pct, fill = course_department))+
    geom_bar(stat = "identity")+
    geom_text(aes(label=mean_growth_pct), position=position_dodge(width=0.9), vjust=-0.25)+
    scale_x_discrete()+
    theme(axis.text.x = element_text(angle = 70, hjust = 1))+
    labs(title = "Average Yearly Enrollment Growth Rate of Top 10 Departments",
         subtitle = "AY 17-18 to AY 19-20",
         fill = "Department")+
    xlab("")+
    ylab("% growth")
```

Slope then confidence in slope

**YO HOW DO I DO THIS**
```{r}
longer_depts_17_to_20 <- depts_17_to20 %>%
  na.omit()%>%
  pivot_longer(values_to = "enrollment",
               names_prefix = "dept_enrollemnts",
               cols = 2:4)


longer_depts_17_to_20 %>%
lm(enrollment~course_department, data = .) %>%
  tidy(conf.int = TRUE)%>%
  select(term, estimate, conf.low, conf.high)%>%
  gt()


```


```{r FASTEST GROWING LINE PLOT}
# USE DEPTS 17 TO 20 TO PIVOT LONGER AND FIRE IN AN ELITE LINE GRAPH
```

FASTEST GROWING BY COURSE COUNT 


```{r 3Y Dept Course Count }
# CREATE DEPT COURSE COUNT 
dept_cc1819 <- ay_1819 %>%
  group_by(course_department)%>%
  summarize(course_count = n())%>%
  filter(course_department != "Medical Sciences")

dept_cc18_to_20 <- dept_cc1819%>%
  full_join(dept_cc1920, by = "course_department", suffix = c("18", "19"))

dept_cc1718 <- ay_1718 %>%
  group_by(course_department)%>%
  summarize(course_count = n())%>%
  filter(course_department != "Medical Sciences")

dept_cc17_to_20 <- dept_cc1718%>%
  full_join(dept_cc18_to_20, by = "course_department")%>%
  rename(course_count17 = course_count)

```

```{r Dept 3Y Course Count Grwowth}

# creating and joining summarized course count tables 

cc_growth_1718 <- dept_cc17_to_20 %>%
  mutate(growth_1718_pct = ((course_count18 - course_count17)/course_count17),
         growth_1718 = (course_count18 - course_count17))

cc_growth_1819 <- dept_cc17_to_20%>%
  mutate(growth_1819_pct = ((course_count19 - course_count18)/course_count18),
         growth_1819 = (course_count19 - course_count18))

cc_growth_17_to_19 <- cc_growth_1718 %>%
  right_join(cc_growth_1819)

# creating mean growth over three years

cc_mean_growth_3y <- cc_growth_17_to_19%>%
  filter(course_department != "Medical Sciences")%>%
  na.omit()%>%
  mutate(mean_growth_cc = (growth_1718 + growth_1819)/2,
            mean_growth_pct_cc = (((growth_1819_pct + growth_1819_pct)/2)*(100)))%>%
  mutate_if(is.numeric, ~round(., 2))
```


```{r}
cc_mean_growth_3y %>%
  arrange(desc(mean_growth_cc))%>%
  slice(1:10)%>%
  ggplot(aes(course_department, mean_growth_cc, fill = course_department))+
    geom_bar(stat = "identity")+
    geom_text(aes(label=mean_growth_cc), position=position_dodge(width=0.9), vjust=-0.25)+
    scale_x_discrete()+
    theme(axis.text.x = element_text(angle = 70, hjust = 1))+
    labs(title = "Average Yearly Course Offering Growth Rate of Top 10 Departments",
         subtitle = "AY 17-18 to AY 19-20",
         fill = "Department")+
    xlab("")+
    ylab("Courses Added")
```

